---
title: "Complexity Doodle"
author: "Andres Kütt"
date: "12/30/2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
library(scales)
library(ggplot2)
library(tidyr)

library(readr)
complexities <- read_delim("~/work/complexity/data/complexities.txt", 
                           "\t", escape_double = FALSE, col_names = FALSE, 
                           col_types = cols(X1 = col_character()), 
                           trim_ws = TRUE)
colnames(complexities) <- c("lbl","month", "year", "name", "C1", "C2", "C3","C")

complexities$date <- as.Date(ymd(paste(complexities$year, complexities$month, 1)))

qc <- read_delim("~/work/complexity/data/q_complexities.txt", 
                           "\t", escape_double = FALSE, col_names = FALSE, 
                           col_types = cols(X1 = col_character()), 
                           trim_ws = TRUE)
colnames(qc) <- c("month", "year", "C1", "C2", "C3","C")

qc$date <- as.Date(ymd(paste(qc$year, qc$month, 1)))


```

## The main thing


```{r main_chart, echo=TRUE, message=FALSE, warning=FALSE}
p <- ggplot(data = qc, aes(x=date, y=C, group = 1)) +
  geom_line() + 
  geom_smooth() + 
  theme_minimal() +
  scale_y_continuous("Keerukus") +
  scale_x_date(date_breaks = "6 months", labels = date_format("%m-%Y"), date_minor_breaks = "1 months", name ="Kuud")+
  theme(axis.text.x = element_text(angle=90))
p
```

### The FB post:

So here it is. How the complexity of Estonian laws has evolved since Q2 2002. I took all the laws in force at 1st of a given quarter, calculated complexity of each law using a formula by prof. de Weck (text complexity was estimated using a morphological method) and then calculated the complexity of the entire set. These are not the final results but the first ones I dare share. Observations:

* It is a strong upwards trend but not a stable one. Not an exponent either, which is good
* Things like EU accession seem to have very little effect although I must enquiry to specialists about the apparent plateau 2008-2011
* It fluctuates quite a bit more than I was expecting. There are a few quite significant drops

### Hypothesis 
#### On the "plateau"
* Economic crysis (several objections)

#### On the drops
* The bugfix hypothesis. People re-write laws, forget to put in references and add those later

### What to do
* Plot all three components on the same figure after converting to 0-1 scale. Differences in growth?
* Show that the individual laws separately get more complex much slower than the set. Thus the need to look at the stucture of the entire thing
* Calculate linear trends of all individual laws (min/max). Do groups emerge? Do scatter plot with min on one and max on other scale

## Individual laws
```{r law_chart, echo=TRUE, message=FALSE, warning=FALSE}
ggplot(data=complexities, aes(x=complexities$date, y=complexities$C)) + 
  geom_line(aes(colour=complexities$name)) + theme(legend.position="none") + scale_y_log10()
```

### Changes in complexity

```{r changes, message=FALSE, warning=FALSE, echo=TRUE}
minmaks <- complexities %>% group_by(name) %>% summarise(i = min(C), x = max(C))
minmaks$g <- minmaks$i/minmaks$x
minmaks$g <- minmaks$i/minmaks$x

ggplot(data = minmaks) + 
  aes(g) + 
  geom_histogram(aes(y=100*..count../sum(..count..))) + 
  theme_minimal()

```

### Descriptive stats

```{r count, message=FALSE, warning=FALSE, echo=TRUE}
mitu <- complexities %>% group_by(date) %>% summarise(mitu = n())
mitu %>% ggplot(aes(x = date, y=mitu)) + 
  geom_line() +
  scale_x_date(date_breaks = "6 months", labels = date_format("%m-%Y"), date_minor_breaks = "1 months", name ="Kuu") + 
  theme_minimal() +
  theme(axis.text.x = element_text(angle=90))
```

### Components of complexity over time
```{r nc, message=FALSE, warning=FALSE, echo=TRUE}
qc$nc1 <- (qc$C1-min(qc$C1))/max(qc$C1)
qc$nc2 <- (qc$C2-min(qc$C2))/max(qc$C2)
qc$nc3 <- (qc$C3-min(qc$C3))/max(qc$C3)
qc$nc <- (qc$C-min(qc$C))/max(qc$C)

qc %>% 
  select(nc1, nc2, nc3, nc, date) %>% 
  gather(type,  value=cmpl, nc1, nc2, nc3, nc, -date) -> nqc
nqc %>%  ggplot(aes(x=date,y=nqc$cmpl, colour = nqc$type)) +
  geom_line()

```

## Narrative
1. Background and research question
1. Scope of work
1. Description of the dataset
1. Key result (upwards trend)
1. Subquestions
    1. Individual components of complexity
    1. Changes in law complexities

## Random notes
Jaanus Tehver FBs:
Päris huvitav on vaadata ka uute terviktekstide vs muutmise seaduste vastuvõtmiste arvu Riigikogu koosseisude kaupa:

Terviktekstid:
VII (1992-1995) 169 
VIII (1995-1999) 155 
IX (1999-2003) 150 
X (2003-2007) 94
XI (2007-2011) 66 
XII (2011-2015) 38

Muutmisseadused: 
VII (1992-1995) 139 
VIII (1995-1999) 345 
IX (1999-2003) 490 
X (2003-2007) 414
XI (2007-2011) 408
XII (2011-2015) 367
"See statistika näitab, et vastuvõetud seaduste (nii uute terviktekstide kui muudatusseaduste) arv vähenes eelmise perioodiga võrreldes kõige tuntavamalt vahetult pärast Eesti liitumist EL-iga. Mis toimus 2008-2011 seaduste keerukusega, siit muidugi ei selgu..."

Tanel Pärnamaa: http://andmeteadus.github.io/2015/2015/05/19/11-seadused/ loetavusindeks